flowers-pilot-analysis

### 2021-08-20 code/aa-flowers-pilot_analysis.Rmd
add nonlang conditions for comparison

# Pre-process
Remove joined_data and uncomment the following block to rerun the preprocessing scripts
```{r process-data, message=FALSE, warning=FALSE}
data_list <- c("pilots")
process_all_data(data_list)
join_all_data()
```

@@

```{r}
#exclude BLOCKS with <12 completed (all 3 participants responded) rounds
#exclude trial runs
# exclude 
rounds_exclude <- d.rounds %>% group_by(gameId, blockNum) %>% tally() %>% filter(n!=18) %>% select(gameId,blockNum)

#exclude cooperative condition
d.rounds.final <-  d.rounds %>% left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
  filter(conditionName != "coopCartel")%>% anti_join(rounds_exclude)

d.raw_chat.final <- d.raw_chat %>% left_join(d.games %>% select(gameId, conditionName, condition)) %>% anti_join(rounds_exclude) %>% write_csv(here("data/processed_data/joined_data/filtered_raw_chat.csv"))
```

@@

```{r}
game_completion <- d.rounds %>% 
  group_by(gameId) %>% 
  summarize(num_rounds=max(trialNum+1)) %>% 
  left_join(d.games%>% ungroup() %>% select(gameId, condition, chat=chatEnabled, gameLength)) %>%
  mutate(gameComplete = ifelse(num_rounds == 24, TRUE, FALSE))

@@

Game counts
```{r}
#pull out pilots
d.games %>% group_by(conditionName, chatEnabled,pilot)%>% tally()
```

@@

# Basic Analyses
```{r}
#get the # of raw points, max adjusted points, game legnth, # of words exchanged
d.round_word_counts <- d.raw_chat.final %>% 
  filter(type == "message") %>%
  full_join(d.rounds.final, c("gameId", "trialNum", "repNum", 
                              "playerId", "numPlayers", "blockNum")) %>%
 # filter(!is.chitchat) %>% 
  mutate(text = gsub('[[:punct:] ]+',' ',text),
         text = str_squish(text),
         utt_length_chars = str_length(text), 
         utt_length_words = str_count(text, "\\W+") + 1) %>% 
  group_by(gameId, blockNum, trialNum, repNum, playerId, numPlayers) %>%
  summarize(text = paste0(text, collapse = ', '),
            total_num_words = sum(utt_length_words),
            total_num_chars = sum(utt_length_chars))

d.max_round <- d.contexts %>%
  group_by(gameId, repNum, blockNum, trialNum) %>%
  #top 3 flowers
  slice_max(utility, n = 3) %>%
  summarise(roundMax = sum(utility))

d.by_round_metrics <- d.rounds.final  %>% 
  left_join(d.round_word_counts) %>%
  group_by(gameId, trialNum, repNum, blockNum) %>%
  summarize(groupPoints = sum(playerUtility, na.rm = T),
            groupNumWords = sum(total_num_words, na.rm = T),
            groupNumChars = sum(total_num_chars, na.rm = T),
            numFlowers = n_distinct(playerResponse)) %>%
  left_join(d.max_round) %>%
  left_join(game_completion %>% select(gameId, gameComplete)) %>%
  left_join(d.games %>% select(-createdAt)) %>%  
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         roundPointsAdjusted = groupPoints/roundMax)

d.by_block_metrics <- d.by_round_metrics %>% 
  group_by(gameId, name, gameComplete, condition, 
           languageCondition, fullCondition, gameLength, blockNum) %>%
  summarize(blockPoints = sum(groupPoints, na.rm = T),
            blockNumWords = sum(groupNumWords, na.rm = T),
            blockNumChars = sum(groupNumChars, na.rm = T),
            blockMaxPoints = sum(roundMax),
            blockNumFlowers = sum(numFlowers)) %>%
  mutate(blockPointsAdjusted = blockPoints/blockMaxPoints,
         meanNumFlowers = blockNumFlowers*3/(12*3))

d.by_game_metrics <- d.by_block_metrics %>%
  #from this point, only look at completed games
    filter(gameComplete) %>%
  #group by game
  group_by(gameId, name, condition, languageCondition, fullCondition, gameLength) %>%
  summarize(gamePoints = sum(blockPoints, na.rm = T),
            gameNumWords = sum(blockNumWords, na.rm = T),
            gameNumChars = sum(blockNumChars, na.rm = T),
            gameMaxPoints = sum(blockMaxPoints),
            gameNumFlowers = sum(blockNumFlowers)) %>%
  mutate(pointsAdjusted = gamePoints/gameMaxPoints,
         meanNumFlowers = gameNumFlowers*3/ (12*4*3))
```

## Simple Distributions
### Points

@@

```{r, fig.height=3, fig.width=5}
d.by_game_metrics%>% 
  ggplot(aes(x = languageCondition, y = gamePoints, color = condition, shape = languageCondition)) + 
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  facet_grid(cols = vars(condition)) + labs(x = "Language Condition", y = "Points Earned by Group", title = "Points Earned by Group") + scale_color_bay() + theme_jmank() + theme(legend.position = "none")

@@

### Time
```{r, fig.height=5, fig.width=3}
d.by_game_metrics %>%
  ggplot(aes(x = languageCondition, y = gameLength/60, 
             color = condition, shape = languageCondition)) + 
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  facet_grid(cols = vars(condition))+ 
  labs(y = "Time (In Minutes)", title = "Game Length", x = "Language Condition")+
  ylim(0, 90)+
  scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
ggsave(here(fig_path, "time_distirbution.png"),  width = 3, height = 5)

@@

### Number of Words
```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = languageCondition, y = gameNumWords, color = condition, shape = languageCondition)) + 
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  facet_grid(cols = vars(condition))+ labs(y = "N. Words", title = "Number of Words Exchanged", x = "Language Condition")+scale_color_bay() + theme_jmank() +  theme(legend.position = "none")

@@

## Covariate
### Time x Points
```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = gameNumWords, y = gameLength/60, color = condition, shape = languageCondition)) + 
  geom_point(alpha = .5) +
  geom_smooth(method=lm)+ 
  labs(x = "Game length (in Minutes)", y = "Points Earned by Group")+
  scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
ggsave(here(fig_path, "time_words_scatter.png"))
```
# All Correlated
```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = gameLength/60, y = gamePoints, shape = languageCondition, color = log(gameNumWords+1))) + 
  geom_point(alpha = .8) +
  geom_smooth(aes(group = condition), method=lm, color = "black")+
  facet_grid(cols = vars(condition)) + 
  labs(x = "Game length (in Minutes)", 
       y = "Points Earned by Group",
       color = "Log Num. Words",
       shape = "Language") + 
  theme_jmank() +  scale_color_steps(low = "#edd746", high = "#6B9555")

@@

```{r}
d.by_game_metrics %>% filter(languageCondition == "Lang") %>%
  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>%
  ggplot(aes(x = gameNumWords, y = gamePoints, color = condition)) + 
  geom_point(alpha = .5) +
  geom_smooth(method=lm)+
  labs(x = "N. Words Exchanged", y = "Group Points (Raw)")+scale_color_bay() + theme_jmank() +  theme(legend.position = "none")

ms <- d.by_game_metrics %>%
  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>% 
  group_by(condition, fullCondition, languageCondition) %>%
  summarise(mean_nw = mean(gameNumWords),
            ci_nw = 1.96 * sd(gameNumWords) / sqrt(length(gameLength)),
            mean_points = mean(gamePoints),
            ci_points = 1.96 * sd(gamePoints) / sqrt(length(gameLength)))

@@

```{r}
ms <- d.by_game_metrics %>%
  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>% 
  group_by(condition, langUse) %>%
  summarise(mean_nw = mean(gameNumWords),
            ci_nw = 1.96 * sd(gameNumWords) / sqrt(length(gameLength)),
            mean_points = mean(gamePoints),
            ci_points = 1.96 * sd(gamePoints) / sqrt(length(gameLength)))

d.by_game_metrics %>%
  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>% 
  ggplot(aes(x = gameNumWords, y = gamePoints, color = condition, shape = langUse)) +
  #geom_smooth(aes(group = condition), method = "lm") +
  geom_point(alpha = .5) +
  geom_pointrange(data = ms, aes(x = mean_nw, y = mean_points, 
                                 ymin = mean_points - ci_points,
                                 ymax = mean_points + ci_points)) +
  geom_errorbarh(data = ms, aes(x = mean_nw, y = mean_points, 
                                 xmin = mean_nw - ci_nw,
                                 xmax = mean_nw + ci_nw), width = 0) + 
  # ylim(0,1500) + 
  labs(x = "Words Echanged", y = "Points Earned") + 
  scale_color_bay() + theme_jmank() +  
  theme(legend.position = "bottom")
```

@@

Including language optional games where they decided not to speak

```{r, fig.height=3, fig.width=5}
d.round_word_counts %>%left_join(d.games %>% 
                                   select(gameId, condition,chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(languageCondition == "Lang") %>%
  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>% 
  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
  ggplot(aes(x = repNum, y = groupNumWords, color = fullCondition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ scale_color_bay() + theme_jmank()
ggsave(here(fig_path, "nwords_over_time.png"))
```

log (n words+1)
```{r, fig.height=3, fig.width=5}
d.round_word_counts %>%left_join(d.games %>% 
                                   select(gameId, condition,chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(languageCondition == "Lang") %>%
  group_by(gameId, condition,block_name, trialNum, repNum) %>% 
  summarize(groupNumWords = sum(total_num_words, na.rm = T)+1) %>%
  ggplot(aes(x = repNum, y = log(groupNumWords), color = condition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name)) + 
  labs(x = "Trial Number", y = "Log (N. Words + 1)")+ scale_color_bay() + theme_jmank() + theme(legend.position = "bottom")

@@

Only games where language was used

```{r, fig.height=3, fig.width=5}
d.round_word_counts %>%left_join(d.games %>% 
                                   select(gameId, condition,chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(languageCondition == "Lang") %>%
  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>% 
  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
  filter(groupNumWords > 0) %>%
  ggplot(aes(x = repNum, y = groupNumWords, color = fullCondition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ scale_color_bay() + theme_jmank()
```

```{r, fig.height=3, fig.width=5}
d.round_word_counts %>%left_join(d.games %>% 
                                   select(gameId, condition,chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(languageCondition == "Lang") %>%
  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>% 
  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
  filter(groupNumWords > 0) %>%
  ggplot(aes(x = repNum, y = log(groupNumWords+1), color = fullCondition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ scale_color_bay() + theme_jmank()
```

log num words

```{r, fig.height=3, fig.width=5}
d.round_word_counts %>%left_join(d.games %>% 
                                   select(gameId, condition, chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(languageCondition == "Lang") %>%
  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>%

  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
  #filter(groupNumWords >1) %>%
  ggplot(aes(x = repNum, y = log(groupNumWords+1), color = fullCondition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  ylim(0, 6) +
  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ scale_color_bay() + theme_jmank()
```
### Accuracy
#### N. Flowers Selected each round
```{r}
d.rounds.final %>% group_by(gameId, repNum, blockNum, trialNum) %>% 
  summarize(num_unique_flowers = n_distinct(playerResponse),
            mean_response_time = mean(time_sec, na.rm = T))  %>%
  left_join(d.games %>% select(gameId, condition,chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(condition != "coopCartel") %>%
  ggplot(aes(x = repNum, y = num_unique_flowers, color = condition, linetype = languageCondition)) + 
  geom_point(alpha = .1) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name), rows = vars(condition)) + labs(x = "Trial Number", y = "N. Unique Flowers")+ scale_color_bay() + theme_jmank()

ggsave(ggsave(here(fig_path, "collisions.png")))
```

```{r}
d.rounds.final %>% group_by(gameId, repNum, blockNum, trialNum) %>% 
  summarize(num_unique_flowers = n_distinct(playerResponse),
            mean_response_time = mean(time_sec, na.rm = T))  %>%
  left_join(d.games %>% select(gameId, condition,chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(condition != "coopCartel") %>%
  ggplot(aes(x = repNum, y = log(mean_response_time), color = condition, linetype = languageCondition)) + 
  geom_point(alpha = .1) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name), rows = vars(languageCondition)) + labs(x = "Trial Number", y = "log Mean response time") + scale_color_bay() + theme_jmank()
```

@@

Are they getting better over games?
```{r}
d.by_round_metrics %>% left_join(d.games %>% 
                                   select(gameId, condition, chatEnabled))%>%
  mutate(blockName = paste0("Block ", blockNum))%>% 
  ggplot(aes(x = log(groupNumWords+1), y = roundPointsAdjusted, 
             color = condition)) + 
  geom_point(alpha = .25) + theme_jmank() + scale_color_bay() + facet_wrap(~blockName) +
  labs(title = "Number of points per trial")
```

>>>>

# Pre-process
Remove joined_data and uncomment the following block to rerun the preprocessing scripts
```{r process-data, message=FALSE, warning=FALSE}
data_list <- c("pilots", "collab_lang_20", "compet_lang_20", "nonlang_pilots", "compet_nonlang_15", "collab_nonlang_15")
process_all_data(data_list)
#join_all_data()
```

@@

```{r}
#exclude BLOCKS with <12 completed (all 3 participants responded) rounds
# exclude incomplete games (not just incomplete rounds)

games_exclude <- d.rounds %>% group_by(gameId) %>% tally() %>% filter(n!=3*24) %>% select(gameId)

#exclude cooperative condition
d.rounds.final <-  d.rounds %>% left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
  filter(conditionName != "coopCartel") %>% anti_join(games_exclude)

d.raw_chat.final <- d.raw_chat %>% 
  left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
  anti_join(games_exclude) %>% 
  write_csv(here("data/processed_data/joined_data/filtered_raw_chat.csv"))
```

@@

```{r}
#how many games are we left with
game_completion <- d.rounds %>% 
  group_by(gameId) %>% 
  summarize(num_rounds=max(trialNum+1)) %>% 
  left_join(d.games%>% ungroup() %>% select(gameId, condition, chat=chatEnabled, gameLength)) %>%
  mutate(gameComplete = ifelse(num_rounds == 24, TRUE, FALSE))

@@

Game counts
```{r}
#pull out pilots
d.rounds.final %>% left_join(d.games %>% select(gameId, chatEnabled)) %>% group_by(condition, chatEnabled) %>% 
  summarize(n_games = n_distinct(gameId)) %>% arrange(condition, chatEnabled)
```

@@

# Basic Analyses

Break out the by-player statistics for each round
```{r player-stats}

#player language per round
d.player_word_counts <- d.raw_chat.final %>% 
  filter(type == "message") %>%
  full_join(d.rounds.final, c("gameId", "trialNum", "repNum", 
                              "playerId", "numPlayers", "blockNum")) %>%
 # filter(!is.chitchat) %>% 
  mutate(text = gsub('[[:punct:] ]+',' ',text),
         text = str_squish(text),
         utt_length_chars = str_length(text), 
         utt_length_words = str_count(text, "\\W+") + 1) %>% 
  group_by(gameId, blockNum, trialNum, repNum, playerId, numPlayers) %>%
  summarize(text = paste0(text, collapse = ', '),
            total_num_words = sum(utt_length_words),
            total_num_chars = sum(utt_length_chars))

# player points per round
d.by_player_metrics <- d.rounds.final  %>% 
  left_join(d.player_word_counts) %>%
  left_join(game_completion %>% 
              select(gameId, gameComplete)) %>%
  left_join(d.games %>% select(-c(createdAt, flowers:scale), condition)) %>%  
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"))

Round info
```{r round-stats}
#get the # of raw points, max adjusted points, game legnth, # of words exchanged
d.max_round <- d.contexts %>%
  group_by(gameId, repNum, blockNum, trialNum) %>%
  #top 3 flowers
  slice_max(utility, n = 3) %>%
  summarise(roundMax = sum(utility))

d.by_round_metrics <- d.by_player_metrics  %>% 
  group_by(gameId, trialNum, repNum, blockNum) %>%
  summarize(groupPoints = sum(playerUtility, na.rm = T),
            groupNumWords = sum(total_num_words, na.rm = T),
            groupNumChars = sum(total_num_chars, na.rm = T),
            numFlowers = n_distinct(playerResponse)) %>%
  left_join(d.max_round) %>%
  left_join(game_completion %>% 
              select(gameId, gameComplete)) %>%
  left_join(d.games %>% select(-c(createdAt, flowers:scale), condition)) %>%  
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         roundPointsAdjusted = groupPoints/roundMax)
```

by block
```{r}
d.by_block_metrics <- d.by_round_metrics %>% 
  group_by(gameId, gameComplete, condition, 
           languageCondition, fullCondition, gameLength, blockNum) %>%
  summarize(blockPoints = sum(groupPoints, na.rm = T),
            blockNumWords = sum(groupNumWords, na.rm = T),
            blockNumChars = sum(groupNumChars, na.rm = T),
            blockMaxPoints = sum(roundMax),
            blockNumFlowers = sum(numFlowers)) %>%
  mutate(blockPointsAdjusted = blockPoints/blockMaxPoints,
         meanNumFlowers = blockNumFlowers*3/(6*3))
```

by game
```{r by-game}
d.by_game_metrics <- d.by_block_metrics %>%
  #from this point, only look at completed games
    filter(gameComplete) %>%
  #group by game
  group_by(gameId, condition, languageCondition, fullCondition, gameLength) %>%
  summarize(gamePoints = sum(blockPoints, na.rm = T),
            gameNumWords = sum(blockNumWords, na.rm = T),
            gameNumChars = sum(blockNumChars, na.rm = T),
            gameMaxPoints = sum(blockMaxPoints),
            gameNumFlowers = sum(blockNumFlowers)) %>%
  mutate(pointsAdjusted = gamePoints/gameMaxPoints,
         meanNumFlowers = gameNumFlowers*3/ (6*4*3))
```

```{r}
d.by_player_metrics %>% group_by(gameId, condition, languageCondition, playerId) %>% 
  summarize(mean_player_utility = mean(playerUtility)) %>%
  group_by(condition, languageCondition) %>% 
  summarize_at("mean_player_utility", funs(mean, min, max, sd))
```
## Simple Distributions
### Points

@@

```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>% 
  ggplot(aes(x = languageCondition, y = gamePoints, color = condition, shape = languageCondition)) + 
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  facet_grid(cols = vars(condition)) + labs(x = "Language Condition", y = "Points Earned by Group", title = "Points Earned by Group") + scale_color_bay() + theme_jmank() + theme(legend.position = "none")

@@

### Time
```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = languageCondition, y = gameLength/60, 
             color = condition, shape = languageCondition)) + 
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  facet_grid(cols = vars(condition))+ 
  labs(y = "Time (In Minutes)", title = "Game Length", x = "Condition")+
  scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
ggsave(here(fig_path, "time_distirbution.png"),  width = 5, height = 3)

@@

### Number of Words
```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = languageCondition, y = gameNumWords, color = condition, shape = languageCondition)) + 
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  facet_grid(cols = vars(condition))+ 
  labs(y = "N. Words", title = "Number of Words Exchanged", x = "Condition")+scale_color_bay() + theme_jmank() +  theme(legend.position = "none")

@@

## Covariate
### Time x Points
```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(y = gamePoints, x = gameLength/60, color = condition, shape = languageCondition)) + 
  geom_point(alpha = .5) +
  geom_smooth(method=lm)+ 
  labs(x = "Game length (in Minutes)", y = "Points Earned by Group")+
  scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
ggsave(here(fig_path, "time_words_scatter.png"))
```
# All Correlated
```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = gameLength/60, y = gamePoints, shape = languageCondition, color = log(gameNumWords+1))) + 
  geom_point(alpha = .8) +
  geom_smooth(aes(group = condition), method=lm, color = "black")+
  facet_grid(cols = vars(condition)) + 
  labs(x = "Game length (in Minutes)", 
       y = "Points Earned by Group",
       color = "Log Num. Words",
       shape = "Language") + 
  theme_jmank() +  scale_color_steps(low = "#edd746", high = "#6B9555")

@@

```{r}
ms <- d.by_game_metrics %>%
  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>% 
  group_by(condition, fullCondition, languageCondition) %>%
  summarise(mean_nw = mean(gameNumWords),
            ci_nw = 1.96 * sd(gameNumWords) / sqrt(length(gameLength)),
            mean_points = mean(gamePoints),
            ci_points = 1.96 * sd(gamePoints) / sqrt(length(gameLength)))

@@

@@

Including language _optional_ games where they decided not to speak

```{r, fig.height=3, fig.width=5}
d.player_word_counts %>%
  left_join(d.games %>% 
              select(gameId, condition,chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(languageCondition == "Lang") %>%
  group_by(gameId, fullCondition, block_name, trialNum, repNum) %>% 
  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
  ggplot(aes(x = repNum, y = groupNumWords, color = fullCondition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ scale_color_bay() + theme_jmank()
ggsave(here(fig_path, "nwords_over_time.png"))
```

log (n words+1)
```{r, fig.height=3, fig.width=5}
d.player_word_counts %>%left_join(d.games %>% 
                                   select(gameId, condition,chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(languageCondition == "Lang") %>%
  group_by(gameId, condition,block_name, trialNum, repNum) %>% 
  summarize(groupNumWords = sum(total_num_words, na.rm = T)+1) %>%
  ggplot(aes(x = repNum, y = log(groupNumWords), color = condition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name)) + 
  labs(x = "Trial Number", y = "Log (N. Words + 1)")+ scale_color_bay() + theme_jmank() + theme(legend.position = "bottom")

@@

Only games where language was _used_

```{r, fig.height=3, fig.width=5}
d.player_word_counts %>%left_join(d.games %>% 
                                   select(gameId, condition,chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(languageCondition == "Lang") %>%
  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>% 
  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
  filter(groupNumWords > 0) %>%
  ggplot(aes(x = repNum, y = groupNumWords, color = fullCondition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ scale_color_bay() + theme_jmank()
```
log num words
```{r, fig.height=3, fig.width=5}
d.player_word_counts %>%left_join(d.games %>% 
                                   select(gameId, condition,chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(languageCondition == "Lang") %>%
  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>% 
  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
  filter(groupNumWords > 0) %>%
  ggplot(aes(x = repNum, y = log(groupNumWords+1), color = fullCondition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "Log N. Words")+ scale_color_bay() + theme_jmank()
```

### Accuracy
#### N. Flowers Selected each round
```{r}
d.by_round_metrics %>% 
  mutate(block_name = paste0("Block ", blockNum)) %>%
  ggplot(aes(x = repNum, y = numFlowers, color = condition, linetype = languageCondition)) + 
  geom_point(alpha = .1) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name), rows = vars(condition)) + labs(x = "Trial Number", y = "N. Unique Flowers")+ scale_color_bay() + theme_jmank()

ggsave(ggsave(here(fig_path, "collisions_conditionfacet.png")))
```
```{r}
d.by_round_metrics %>% 
  mutate(block_name = paste0("Block ", blockNum)) %>%
  ggplot(aes(x = repNum, y = numFlowers, color = condition, linetype = languageCondition)) + 
  geom_point(alpha = .1) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name), rows = vars(languageCondition)) + labs(x = "Trial Number", y = "N. Unique Flowers")+ scale_color_bay() + theme_jmank()

ggsave(ggsave(here(fig_path, "collisions_language_facet.png")))
```

Response Time
```{r}
d.rounds.final %>% group_by(gameId, repNum, blockNum, trialNum) %>% 
  summarize(mean_response_time = mean(time_sec, na.rm = T))  %>%
  left_join(d.games %>% select(gameId, condition,chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(condition != "coopCartel") %>%
  ggplot(aes(x = repNum, y = log(mean_response_time), color = condition, linetype = languageCondition)) + 
  geom_point(alpha = .1) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name), rows = vars(languageCondition)) + labs(x = "Trial Number", y = "log Mean response time") + scale_color_bay() + theme_jmank()
```

@@

Are they getting better over games?
```{r}
d.by_round_metrics %>%
  mutate(blockName = paste0("Block ", blockNum))%>% 
  ggplot(aes(x = groupNumWords, y = roundPointsAdjusted, 
             color = condition)) + 
  geom_point(alpha = .25) + theme_jmank() + scale_color_bay() + facet_wrap(~blockName) +
  labs(title = "Number of points per trial")
```
### Over/underpowered participants

```{r}
d.point_propotions <- d.by_player_metrics %>% 
  #points for entire game
  group_by(condition, languageCondition, gameId, playerId) %>% 
  summarize(player_total = sum(playerUtility)) %>% 
  group_by(condition, gameId) %>% mutate(game_total = sum(player_total)) %>% ungroup() %>%
  mutate(player_proportion = player_total/game_total)

#min
min_prop <- d.point_propotions %>% 
  group_by(condition, languageCondition, gameId) %>% 
  #lowest earned
  slice_min(order_by = player_proportion, n = 1) %>% 
  filter(condition == "Competitive") %>% select(-condition) %>% 
  mutate(player_placement = "Lowest Earning")

#max
max_prop <- d.point_propotions %>% 
  group_by(condition, languageCondition, gameId) %>% 
  slice_max(order_by = player_proportion, n = 1) %>% 
  filter(condition == "Competitive") %>% 
  select(-condition) %>% 
  mutate(player_placement = "Highest Earning")

rbind(min_prop, max_prop) %>%
  ggplot(aes(x = languageCondition, y = player_proportion, color = condition)) + 
  geom_hline(yintercept=.33, linetype = "dashed", alpha = .5)+
  geom_jitter(width = .1, alpha = .5) +
  facet_grid(cols = vars(player_placement)) + 
  stat_summary(fun.data = "mean_cl_boot")+
  labs(x = "Participant Placement", y = "Points earned by participant in place") + scale_color_bay() + theme_jmank() + theme(legend.position = "none")

```
### Entropy

```{r}
entropy_func <- function(p1, p2, p3){
  return((p1 * log(p1)) + (p2 * log(p2)) + (p3 * log(p3)))
  #return((entropy_list[1] * log(entropy_list[1])) + (entropy_list[2] * log(entropy_list[2])) + (entropy_list[3] * log(entropy_list[3])))
}
d.game_entropies <- d.point_propotions %>% select(condition, languageCondition, gameId, player_proportion) %>% 
  mutate(p_num = rep(c("p1","p2","p3"), nrow(d.point_propotions)/3)) %>% 
  pivot_wider(names_from = p_num, values_from = player_proportion) %>% 
  mutate(entropy = 0-entropy_func(p1, p2, p3)) %>% select(-c(p1,p2,p3))

d.game_entropies %>% filter(condition == "Competitive") %>% 
  ggplot(aes(x = languageCondition, y = entropy)) + 
  geom_hline(yintercept=(0-entropy_func(1/3, 1/3, 1/3)), linetype = "dashed", alpha = .5)+
  #minimum is tricky? this is 12, 1, and 2 which is the least fair a game _could possible_ be, but isnt adjusted per game
  #geom_hline(yintercept=(0-entropy_func(12/15, 1/15, 2/15)), linetype = "dashed", alpha = .5)+
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  labs(x = "Condition", y = "Entropy of Player Scores (by game)") + scale_color_bay() + theme_jmank() + theme(legend.position = "none")
```
